Curation of NYC 311 Request Open Data

### 2023-06-06 talk/slides.Rmd
reduced default r chunk font size and minor edits to slides

```{r, echo = FALSE}
library(stringr)
source("../code/functions.R")
## Load the 311 SR data file
```

@@

## Invalid Data: Created/Closed Date
```{r date}
t1 <- strptime(d311$closed_date, format = '%m/%d/%Y %I:%M:%S %p')
t0 <- strptime(d311$created_date, format = '%m/%d/%Y %I:%M:%S %p')
tt <- as.numeric(difftime(t1, t0, units =  "secs"))
badTime <- subset(d311, tt <= 0,
                  select = c("created_date", "closed_date", "agency"))
nrow(badTime)
sum(tt == 0, na.rm = TRUE)
summary(tt[tt < 0 & (!is.na(tt))])
table(badTime$agency)
```

@@

## 'borough' and 'borough_boundaries' are almost identical
nonMatchingBorough <- subset(
    d311, borough_boundaries !=
          as.integer(str_replace_all(borough,
                                     c("STATEN ISLAND" = "1",
                                       "BROOKLYN"      = "2",
                                       "QUEENS"        = "3",
                                       "MANHATTAN"     = "4",
                                       "BRONX"         = "5",
                                       "UNSPECIFIED"   = NA))),
    select = c("borough", "borough_boundaries", "agency"))
nrow(nonMatchingBorough)
```

## Redundant Data: Latitude, Longitude, and Location
```{r location}
head(d311[, c("latitude", "longitude", "location")])
```

@@

## Redundant Data: Cross Street and Intersection Street (cont'd)
```{r extra_space}
## after removing extra spaces, most of them are the same
dupXStreets <-subset(
    d311, gsub(" ", "", cross_street_1) ==
          gsub(" ", "", intersection_street_1),
    select = c("cross_street_1", "intersection_street_1", "agency"))
head(dupXStreets)
table(dupXStreets$agency)
head(subset(dupXStreets, cross_street_1 != intersection_street_1))
```

## Less-Efficient Storage: Borough, Agency
```{r storage}
## code borough by integer
object.size(d311$borough_boundaries)
object.size(d311$borough)

@@

## Suggestions on Data Curation

+ Validity check 
    - date (logical check)
    - descriptive fields (drop-down menu)
    - legitimate values
+ Efficiency
    - Eliminate redundant columns
    - Reduce storage with no information loss
+ Feedback: open review for open data

>>>>

```{r setup, include = FALSE, echo = FALSE}
library(stringr)
source("../code/functions.R")

def.chunk.hook  <- knitr::knit_hooks$get("chunk")
knitr::knit_hooks$set(chunk = function(x, options) {
  x <- def.chunk.hook(x, options)
  paste0("\n \\", "small","\n\n", x, "\n\n \\normalsize")
})
## Load the 311 SR data file
```

@@

## Invalid Data: Created/Closed Date
```{r date}
t1 <- strptime(d311$closed_date, format = '%m/%d/%Y %I:%M:%S %p')
t0 <- strptime(d311$created_date, format = '%m/%d/%Y %I:%M:%S %p')
tt <- as.numeric(difftime(t1, t0, units =  "secs"))
badTime <- subset(d311, tt <= 0,
                  select = c("created_date", "closed_date", "agency"))

## exactly the same created and closed date down to the second; a lot!
sum(tt == 0, na.rm = TRUE)
table(subset(badTime, created_date == closed_date, select = "agency"))
```

## Bad Closed Date
```{r baddate}
## quite some negative response times
table(subset(badTime, created_date != closed_date, select = "agency"))

summary(tt[tt < 0 & (!is.na(tt))])
```

@@

## 'borough' and 'borough_boundaries' are almost identical
nonMatchingBorough <- subset(
    d311, borough_boundaries !=
          as.integer(str_replace_all(
              borough, c("STATEN ISLAND" = "1",  "BROOKLYN"      = "2",
                         "QUEENS"        = "3",  "MANHATTAN"     = "4",
                         "BRONX"         = "5",  "Unspecified"   = NA))),
    select = c("borough", "borough_boundaries", "agency"))
nrow(nonMatchingBorough)
```

## Redundant Data: Latitude, Longitude, and Location
```{r location}
print(head(d311[, c("latitude", "longitude", "location")]), digits = 16)
```

@@

##  Cross Street and Intersection Street (cont'd)
```{r extra_space}
## after removing extra spaces, most of them are the same
dupXStreets <-subset(
    d311, gsub(" ", "", cross_street_1) ==
          gsub(" ", "", intersection_street_1),
    select = c("cross_street_1", "intersection_street_1", "agency"))
head(dupXStreets)
```

##  Cross Street and Intersection Street (cont'd) 
```{r dup_streets}
table(dupXStreets$agency)
head(subset(dupXStreets, cross_street_1 != intersection_street_1))
```

## Less-Efficient Storage: Download File Size
```{r storage}
## code borough by integer
object.size(d311$borough_boundaries)
object.size(d311$borough)

@@

## Suggestions on Open Data Curation

+ Validity check 
    - date (logical check)
    - legitimate values 
	- descriptive fields (drop-down menu, possibly multi-step)
    - missing value: coding; validity
    - computed fields in the dictionary
+ Efficiency
    - Eliminate redundant columns
    - Reduce storage with no information loss
+ Feedback: 
    - open review for open data
	- interaction with the maintenance team
